---
title: "Toronto Rapters Shot Chart From 2000-2020"
author: "Ben Bronoski"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(ggimage)
library(ggpubr)
library(png)
library(gganimate)
library(ggplot2)

stats = read.csv("Totals.csv")

#Filtered Y axis data to only include shots that were taken from the baseline to halfcourt. Some years have data with missed, full court shots that skewed some graphs
stats = stats %>% filter(stats$locationY<300)
bball = readPNG("Halfcourt.png")
```


## Beginning of the Century

- Very diverse shot taking
- Emphasis on taking shot that was designed

```{r, warning=FALSE}
stats2001 = subset(stats, stats$yearSeason == 2001)
ggplot(data = stats2001, aes(stats2001$locationX, stats2001$locationY, color = stats2001$typeEvent)) +background_image(bball) + geom_point() + scale_color_manual(values = c("Made Shot" = "Red", "Missed Shot" = "Black")) + labs(color = "Shot Outcome", x = NULL, y = NULL) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_y_discrete(labels = NULL, breaks = NULL) + ggtitle("Shot Chart for 1999-2000")
```


## Late 2000's to Early 2010's

- Mid range shooting becomes slightly more prominent
- Large amount of shots still taken within a system

```{r}
stats2005 = subset(stats, stats$yearSeason == 2005)
ggplot(data = stats2005, aes(stats2005$locationX, stats2005$locationY, color = stats2005$typeEvent)) +background_image(bball) + geom_point() + scale_color_manual(values = c("Made Shot" = "Red", "Missed Shot" = "Black")) + labs(color = "Shot Outcome", x = NULL, y = NULL) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_y_discrete(labels = NULL, breaks = NULL) + ggtitle("Shot Chart for 2004-2005")
```


## Late 2000's to Early 2010's cont.

- Clear shot taking patterns are starting to emerge between the post and 3-point line

```{r, warning=FALSE}
stats2010 = subset(stats, stats$yearSeason == 2010)
ggplot(data = stats2010, aes(stats2010$locationX, stats2010$locationY, color = stats2010$typeEvent)) +background_image(bball) + geom_point() + scale_color_manual(values = c("Made Shot" = "Red", "Missed Shot" = "Black")) + labs(color = "Shot Outcome", x = NULL, y = NULL) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_y_discrete(labels = NULL, breaks = NULL) + ggtitle("Shot Chart for 2004-2005")
```

## Late 2010's to 2020

- Start to see the true emergence of the 3-point shot
- Iso ball begins to take shape as shots are focused either in the post or 3-point range

```{r, warning=FALSE}
stats2015 = subset(stats, stats$yearSeason == 2015)
ggplot(data = stats2015, aes(stats2015$locationX, stats2015$locationY, color = stats2015$typeEvent)) +background_image(bball) + geom_point() + scale_color_manual(values = c("Made Shot" = "Red", "Missed Shot" = "Black")) + labs(color = "Shot Outcome", x = NULL, y = NULL) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_y_discrete(labels = NULL, breaks = NULL) + ggtitle("Shot Chart for 2014-2015")
```

## Late 2010's to 2020 cont.

- 3-point shot completely takes over
- Almost all shots are dunks or 3-pointers

```{r, warning=FALSE}
stats2020 = subset(stats, stats$yearSeason == 2020)
ggplot(data = stats2020, aes(stats2020$locationX, stats2020$locationY, color = stats2020$typeEvent)) +background_image(bball) + geom_point() + scale_color_manual(values = c("Made Shot" = "Red", "Missed Shot" = "Black")) + labs(color = "Shot Outcome", x = NULL, y = NULL) + scale_x_discrete(labels = NULL, breaks = NULL) + scale_y_discrete(labels = NULL, breaks = NULL) + ggtitle("Shot Chart for 2019-2020")
```

## Final Thoughts

- The transition of shot selection over time shows players are getting better at longer range shots
- Defenses are pushing players to take longer shots
- More teams are willing to give up a dunk than try to defend down low

## Extra Thoughts Overall

- There is a huge emphasis on the 3-pointer because of guys like Steph Curry
- Defenses are built to attempt to stop the 3 but as the graphs show, the players are just learning to shoot from further and further away
- Most of the shots taken around the rim are due to transition dunks, or tip ins
- This data does not count shots taken from the free throw line, only field goals
- It would be interesting to see if a defense could be built to trick guys into stepping forward across the 3-point line to attempt more long 2's since those seem to be missed a fair amount